---
title: "PA1_template"
author: "Aleena Jacob"
date: "8/6/2021"
output: html_document
---

# Reproducible Research: Peer Assessment 1



## Load and process the data
##### 1. Load the data 

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
library(ggplot2)
library(scales)
library(Hmisc)
```


```{r, results='markup', warning=TRUE, message=TRUE}
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
activityData <- read.csv('activity.csv')
```
##### 2. transforming the interval data

```{r}
#activityData$interval <- strptime(gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", activityData$interval), format='%H:%M')
```

-----

## Mean total number of steps taken per day.

```{r}
stepsByDay <- tapply(activityData$steps, activityData$date, sum, na.rm=TRUE)
```

##### 1. Histogram of the total number of steps taken each day

```{r}
qplot(stepsByDay, xlab='Total steps per day', ylab='Frequency using binwith 500', binwidth=500)
```



##### 2. Mean and median of the total number of steps taken per day

```{r}
meanOfStepsPerDay <- mean(stepsByDay)
medianOfStepsPerDay <- median(stepsByDay)
```
* Mean: `r meanOfStepsPerDay` 
* Median: `r medianOfStepsPerDay`

-----

## Average daily activity pattern?

```{r}
averageStepsPerTime <- aggregate(x=list(meanSteps=activityData$steps), by=list(interval=activityData$interval), FUN=mean, na.rm=TRUE)
```

##### 1. Time series plot

```{r}
ggplot(data=averageStepsPerTime, aes(x=interval, y=meanSteps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken") 
```
##### 2. The 5-minute interval on average across all the days in the dataset that contains the maximum number of steps?

```{r}
mostSteps <- which.max(averageStepsPerTime$meanSteps)
MostSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averageStepsPerTimeBlock[mostSteps,'interval'])
```

* Most Steps at: `r MostSteps`

----

## Imputing missing values
##### 1. The total number of missing values in the dataset 

```{r}
numOfMissingValues <- length(which(is.na(activityData$steps)))
```

* Number of missing values: `r numOfMissingValues`

##### 2. Devise a strategy for filling in all of the missing values in the dataset.
##### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
activityDataImputed <- activityData
activityDataImputed$steps <- impute(activityData$steps, fun=mean)
```


##### 4. Histogram of the total number of steps taken each day 

```{r}
stepsByDayImputed <- tapply(activityDataImputed$steps, activityDataImputed$date, sum)
qplot(stepsByDayImputed, xlab='Total steps per day (Imputed)', ylab='Frequency using binwith 500', binwidth=500)
```


 

##### ... and Calculate and report the mean and median total number of steps taken per day. 

```{r}
stepsByDayMeanImputed <- mean(stepsByDayImputed)
stepsByDayMedianImputed <- median(stepsByDayImputed)
```
* Mean (Imputed): `r stepsByDayMeanImputed`
* Median (Imputed):  `r stepsByDayMedianImputed`


----

## Are there differences in activity patterns between weekdays and weekends?
##### 1. Create a new factor variable in the dataset with two levels weekday and weekend indicating whether a given date is a weekday or weekend day.


```{r}
activityDataImputed$dateType <-  ifelse(as.POSIXlt(activityDataImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

##### 2. Make a panel plot containing a time series plot


```{r}
averagedActivityDataImputed <- aggregate(steps ~ interval + dateType, data=activityDataImputed, mean)
ggplot(averagedActivityDataImputed, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```
